I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-17 Paper/Paper_noProof.tex

To mesure the accuracy, we measure the uncertainty of all nodes in a layer:
the uncertainty of a node is the range between its computed lower and upper bound. 
We then average the uncerainty among all the nodes of the layer.
Formally, for a node $n$ with bounds $[\alpha,\beta]$, its uncertainty $unc(a) = \beta - \alpha$, and the average uncertainty of a layer $l$ is $\dfrac{\sum_{a\in l} unc(a)}{|l|}.$

@@

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer $2$) only is never a good choice. 
When picking in both layer $1$ and $2$, choosing $50$ out of $116$ unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inacurate computations ($0.059$ vs $0.574$), while it only reduces the inacuracy by $65\%$ when choosing nodes randomly. Notice that this is direct inacuracy, applying this layer after layer will even further reduce inacuracy created by accumulation of inacuraccies.

Notice that it is likely that selecting ReLU nodes up to 3 layers earlier (for nodes in layer $4+$) could further improve the accuracy, but the computational trade off is unclear and we thus didnot experiment.

@@

	\item We do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI} which do not present favorable trade offs compared with $\beta$-CROWN. Results can be found in \cite{crown} and are performed in the same conditions so can be compared.
	\item Finally, we run several version of Compensate($K$). First, we have a fixed choice for $K$ ($K=60$ for fast and $K=200$ for slow), with a fall back to a lower $K$ ($K=50$ for fast and $K=60$ for slow) if the bounds found is too far from the solution found (MIPGap $>0.5$ in Gurobi). We run (our implementation of) DeepPoly first, then the fast version, before running slow on the remaining Images.
	Further, we also experiment with an {\em adaptative} implementation, which lowers number $K$ adaptively between layers to adapt to the growing number $N$ of linear variables, which has an impact on the overall $\MILP(N,K)$ runtime. This is particularly useful for deeper network where $N$ is very different between the first and the last layer. Again, we run first an aggressive version of fast, before going to a more concervative slow version to treat the remaining images. All versions compute bounds for 20 neurons of the same layer in parallel, no GPU used.
\end{itemize}

In Table \ref{tab:example}, we report the accuracy and runtime of these different methods.
DeepPoly/CROWN is by far the fastest, but also the most inacurate, certifying robustness for less than $30\%$ of the images, whereas there are at least $82\%$ of the images for which no attacked is found (attacks computed by \cite{attack} in $\beta$-CROWN).
This is because these DNNs, altough of reasonable size, are hard to verify (they are trained in a natural way, with high compensation strength, see Section \ref{Sec.comp}).

Compared with PRIMA (using refinement of the first few layers by exact MILP), the overall method is quite similar, looking at all the nodes one by one from the begining of the network till the end. Our method compare favorably, both in terms of time and accuracy: our fastest setting is always faster than PRIMA, sometimes by a large margin ($>3$ times faster on $8 \times 100$ while verifying $10\%$ more images), and also more accurate (except our fast adaptative method for $5 \times 200$ which is quite pointless on this shallow DNN - the fixed method is both faster and more accurate than PRIMA), sometimes by a large margin (by $14 \%$  for $5 \times 100$ while being $30\%$ faster). Further, unlike PRIMA, we do not resort to a GPU, and our purely Python implementation is not optimal. This is very promising for this idea of compensation, which indirectly treats dependencies between the variables, which PRIMA represents explicitly.

@@

Compare with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), 
the picture is more balanced, because the fundamentals of both algorithms are different. BaB 
focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. 
Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence our algorithm will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized
$\beta$-CROWN: our implementation is 2 to 5 times slower than $\beta$-CROWN, while we are not using GPU, unlike $\beta$-CROWN. Things are more interesting on the accuracy front: on shallower DNNs ($5 \times 100$ and $5 \times 100$ with 5 hidden layers), the refinement with MILP treats all but the 2 last layers, 
leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accuate than using compensation: we are within $1.6\%$ of the accuracy. On deeper DNNs ($8 \times 100$ and $8 \times 100$ with 8 hidden layers), the refinement with MILP cannot treat the 4 last layers, and BaB has more ReLU to branch on. Therefore, Compensate($K$) is more accurate than $\beta$-CROWN, 
closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $12\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than Compensate($K$) (see next Section).

>>>>

To measure the accuracy, we measure the uncertainty of all nodes in a layer:
the uncertainty of a node is the range between its computed lower and upper bound. 
We then average the uncertainty among all the nodes of the layer.
Formally, for a node $n$ with bounds $[\alpha,\beta]$, its uncertainty $unc(a) = \beta - \alpha$, and the average uncertainty of a layer $l$ is $\dfrac{\sum_{a\in l} unc(a)}{|l|}.$

@@

We report in Table \ref{tab:example1} the average uncertainty of $\MILP_Z$ following the choice of the $K$ heaviest compensating ReLU nodes in $Z$, vs choosing $K$ nodes randomly from layers $1$ and $2$. 
We compare choosing nodes in the second layer only with choosing nodes in both the first and second layer.
Choosing ReLU nodes in the previous layer (layer $2$) only is never a good choice. 
When picking in both layer $1$ and $2$, choosing $50$ out of $116$ unstable ReLU nodes allows to reduce by $90\%$ the average uncertainty due to inaccurate computations ($0.059$ vs $0.574$), while it only reduces the inaccuracy by $65\%$ when choosing nodes randomly. Notice that this is direct inaccuracy, applying this layer after layer will even further reduce inaccuracy created by accumulation of inaccuracies.

Notice that it is likely that selecting ReLU nodes up to 3 layers earlier (for nodes in layer $4+$) could further improve the accuracy, but the computational trade off is unclear and we thus did not experiment.

@@

	\item We do not report results from kPoly \cite{kpoly}, OptC2V \cite{optC2V}, 
	SDPFO \cite{SDPFI} which do not present favorable trade offs compared with $\beta$-CROWN. Results can be found in \cite{crown} and are performed in the same conditions so can be compared.
	\item Finally, we run several version of Compensate($K$). First, we have a fixed choice for $K$ ($K=60$ for fast and $K=200$ for slow), with a fall back to a lower $K$ ($K=50$ for fast and $K=60$ for slow) if the bounds found is too far from the solution found (MIPGap $>0.5$ in Gurobi). We run (our implementation of) DeepPoly first, then the fast version, before running slow on the remaining Images.
	Further, we also experiment with an {\em adaptative} implementation, which lowers number $K$ adaptively between layers to adapt to the growing number $N$ of linear variables, which has an impact on the overall $\MILP(N,K)$ runtime. This is particularly useful for deeper network where $N$ is very different between the first and the last layer. Again, we run first an aggressive version of fast, before going to a more conservative slow version to treat the remaining images. All versions compute bounds for 20 neurons of the same layer in parallel, no GPU used.
\end{itemize}

In Table \ref{tab:example}, we report the accuracy and runtime of these different methods.
DeepPoly/CROWN is by far the fastest, but also the most inaccurate, certifying robustness for less than $30\%$ of the images, whereas there are at least $82\%$ of the images for which no attacked is found (attacks computed by \cite{attack} in $\beta$-CROWN).
This is because these DNNs, although of reasonable size, are hard to verify (they are trained in a natural way, with high compensation strength, see Section \ref{Sec.comp}).

Compared with PRIMA (using refinement of the first few layers by exact MILP), the overall method is quite similar, looking at all the nodes one by one from the beginning of the network till the end. Our method compare favorably, both in terms of time and accuracy: our fastest setting is always faster than PRIMA, sometimes by a large margin ($>3$ times faster on $8 \times 100$ while verifying $10\%$ more images), and also more accurate (except our fast adaptative method for $5 \times 200$ which is quite pointless on this shallow DNN - the fixed method is both faster and more accurate than PRIMA), sometimes by a large margin (by $14 \%$  for $5 \times 100$ while being $30\%$ faster). Further, unlike PRIMA, we do not resort to a GPU, and our purely Python implementation is not optimal. This is very promising for this idea of compensation, which indirectly treats dependencies between the variables, which PRIMA represents explicitly.

@@

Compare with $\beta$-CROWN (also using a refinement on the first few layers using exact MILP), 
the picture is more balanced, because the fundamentals of both algorithms are different. BaB 
focuses on the output neuron, computing bounds for it, and refining the different ReLU as necessary. 
Instead, we consider every node one by one from the input layer till the final layer, with the same accuracy. Hence our algorithm will spend a lot of time on nodes which are not necessary for the verification. Therefore, our naive implementation is not as efficient as the highly optimized
$\beta$-CROWN: our implementation is 2 to 5 times slower than $\beta$-CROWN, while we are not using GPU, unlike $\beta$-CROWN. Things are more interesting on the accuracy front: on shallower DNNs ($5 \times 100$ and $5 \times 100$ with 5 hidden layers), the refinement with MILP treats all but the 2 last layers, 
leaving BaB with few ReLU nodes to branch on, and $\beta$-CROWN (refined BaB) is slightly more accurate than using compensation: we are within $1.6\%$ of the accuracy. On deeper DNNs ($8 \times 100$ and $8 \times 100$ with 8 hidden layers), the refinement with MILP cannot treat the 4 last layers, and BaB has more ReLU to branch on. Therefore, Compensate($K$) is more accurate than $\beta$-CROWN, 
closing the gap with the upper bound from $20\%$ to $16.2\%$ ($8 \times 100$)
and from $17.6\%$ to $12\%$ ($8 \times 200$). This is very promising for the compensation idea, which could be used in many different ways than Compensate($K$) (see next Section).

### 2024-01-17 Paper/proofs.tex

		\begin{lemma}
			Let $L,L'$ be two consecutive layers.
			Assume for all node $m$ of $L$ that the lower and upper bounds for $m$ used by box abstraction equals to the exact lower and upper bounds of $m$ (as expressed above), i.e.
			$\bar{f}(m)=max(m)$ and $\underline{f}(m)=min(m)$.

			Then Box abstraction computes the exact lower and upper bounds for every node $n$ of layer $L'$, ie $\bar{f}(n)=max(n)$ and $\underline{f}(n)=min(n)$.
		\end{lemma}

@@

				\begin{theorem}

@@

				From Lemma \ref{lem:main}, we already have that for the unique output node $z$, $$\UB(z)=\max(z).$$ This finish the proof of Theorem \ref{th2}.

>>>>

		\begin{lemma}
			Let $L,L'$ be two consecutive layers.
			Assume for all node $m$ of $L$ that the lower and upper bounds for $m$ used by box abstraction equals to the exact lower and upper bounds of $m$ (as expressed above), i.e.
			$\bar{f}(m)=\max(m)$ and $\underline{f}(m)=\min(m)$.

			Then Box abstraction computes the exact lower and upper bounds for every node $n$ of layer $L'$, ie $\bar{f}(n)=\max(n)$ and $\underline{f}(n)=\min(n)$.
		\end{lemma}

@@

				\begin{theorem} \label{thm:2}

@@

				From Lemma \ref{lem:main}, we already have that for the unique output node $z$, $$\UB(z)=\max(z).$$ This finish the proof of Theorem \ref{thm:2} and hence Theorem \ref{th2}.
